\documentclass[onecolumn]{aastex62}


\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\usepackage{listings}
\usepackage{amsmath}
\usepackage{physics}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}

\begin{document}

\title{\Large Milestone II:\\Studying the recombination history of the universe.}


\author{Håkon Tansem}

\section{Introduction} \label{sec:intro}
This is the third milestone in a four milestone project where the end goal is to compute the CMB power spectrum. In the previous two milestones we have solved for the background evolution of the universe(). In this third milestone we will study how perturbations has evolved throughout cosmic history. We will start with the initial conditions as predicted by inflation and solve the equation given by the Boltzmann and Einstein equations ( litt ufullstendig foreløpig).
\section{Method} \label{sec:method}
In this model we assume a few simplifications. We neglect photon polarization and ignore neutrinos as to reduce the number of equations. We will therefore only study the evolution of the density parameter $\delta$ and velocity parameter $v$ for baryons and dark matter, the perturbation to the photon distribution $\theta_{\ell}$, where we include integer multipoles $\ell\in[0, 7]$, as well as the metric perturbations $\Phi$ and $\Psi$ all as functions of $x=log(a)$. When studying the perturbations we will work in fourier space where each mode is denoted by the magnitude of its wavevector $k$. When integrating the perturbations we have two main regimes to take into account. In early times, as we calculated in milestone II, we saw that the optical depth of the universe $\tau\gg1$ due to all electrons being free. This means that at any given point in space the mean free path for a photon is very small. This gives very little anisotropy as only perturbations very close by can be observed. This suppresses all higher multipoles and washes out their gradients making only $\ell=0$ and $\ell=1$ the nonnegligible moments. This makes photons behave like a fluid described by the two variables density and velocity which in our case is represented by $\theta_0$ and $\theta_1$ respectively \citep[p. 225]{Dodelson:1282338}. As criterions for deciding wether we are in the tight coupling regime we check wether we are past the onset of recombination, at which time we approximate by evualuating the free electron fraction $X_e$ and seeing wether it is smaller than $0.99$, or if the criterion $\left|\frac{d\tau}{dx}\right| < 10 \cdot \text{min}(1, \frac{ck}{\mathcal{H}})$ is met. That is if $X_e < 0.99$ or if the criterion is met, we assume that we are no longer in the tight coupling regime.\\\indent

When solving the equations we start by setting the initial conditions given by inflation as
\begin{align}
    \Psi &= -\frac{2}{3}\\
    \Phi &= -\Psi \\
    \delta_{\rm CDM} &= \delta_b = -\frac{3}{2} \Psi \\
    v_{\rm CDM} &= v_b = -\frac{ck}{2\mathcal{H}} \Psi\\
    \Theta_0 &= -\frac{1}{2} \Psi \\
    \Theta_1 &= +\frac{ck}{6\mathcal{H}}\Psi \\
    \Theta_2 &= -\frac{20ck}{45\mathcal{H}\tau^\prime} \Theta_1,\label{eq:t2init}\\
    \Theta_\ell &= -\frac{\ell}{2\ell+1} \frac{ck}{\mathcal{H}\tau^\prime} \Theta_{\ell-1}.\label{eq:tellinit}\\
\end{align}
All equations in this paper are given by \cite{WintherIII:2020} unless otherwise stated. The full set of equations for the photon multipoles are given as
\begin{align}
    \Theta^\prime_0 &= -\frac{ck}{\mathcal{H}} \Theta_1 - \Phi^\prime,,\label{eq:tp1} \\
    \Theta^\prime_1 &=  \frac{ck}{3\mathcal{H}} \Theta_0 - \frac{2ck}{3\mathcal{H}}\Theta_2 +
    \frac{ck}{3\mathcal{H}}\Psi + \tau^\prime\left[\Theta_1 + \frac{1}{3}v_b\right], \\
    \Theta^\prime_\ell &= \frac{\ell ck}{(2\ell+1)\mathcal{H}}\Theta_{\ell-1} - \frac{(\ell+1)ck}{(2\ell+1)\mathcal{H}}
    \Theta_{\ell+1} + \tau^\prime\left[\Theta_\ell - \frac{1}{10}\theta_2
    \delta_{\ell,2}\right], \quad\quad 2 \le \ell < \ell_{\textrm{max}} \\
    \Theta_{\ell}^\prime &= \frac{ck}{\mathcal{H}}
    \Theta_{\ell-1}-c\frac{\ell+1}{\mathcal{H}\eta(x)}\Theta_\ell+\tau^\prime\Theta_\ell,
    \quad\quad \ell = \ell_{\textrm{max}}.\\
\end{align}
where $\delta_{\ell,2}$ is the kroenecker-delta and $\eta(x)$ is the conformal time computed in milestone I. For cold dark matter and baryons as well as the metric perturbations, we have
\begin{align}
    \delta_{\rm CDM}^\prime &= \frac{ck}{\mathcal{H}} v_{\rm CDM} - 3\Phi^\prime \\
    v_{\rm CDM}^\prime &= -v_{\rm CDM} -\frac{ck}{\mathcal{H}} \Psi \\
    \delta_b^\prime &= \frac{ck}{\mathcal{H}}v_b -3\Phi^\prime \\
    v_b^\prime &= -v_b - \frac{ck}{\mathcal{H}}\Psi + \tau^\prime R(3\Theta_1 + v_b) \\
    \Phi^\prime &= \Psi - \frac{c^2k^2}{3\mathcal{H}^2} \Phi + \frac{H_0^2}{2\mathcal{H}^2}
    \left[\Omega_{\rm CDM} e^{-x} \delta_{\rm CDM} + \Omega_{b,0} e^{-x} \delta_b + 4\Omega_{r,0}
    a^{-2}\Theta_0\right] \\
    \Psi &= -\Phi - \frac{12H_0^2}{c^2k^2e^{2x}}\Omega_{r,0}\Theta_2, \\
\end{align}
where $R = \frac{4\Omega_{r,O}}{3\Omega_{b,0}e^x}$. These equations apply as they are written when we have exited the tight coupling regime. In the tight coupling regime, as described earlier, multipoles larger than $\ell=1$ have their gradients washed out. This gives the following changes to the set of equations 
\begin{align}
    q &= \frac{-[(1-R)\tau^\prime + (1+R)\tau^{\prime\prime}](3\Theta_1+v_b) -
    \frac{ck}{\mathcal{H}}\Psi + (1-\frac{\mathcal{H}^\prime}{\mathcal{H}})\frac{ck}{\mathcal{H}}(-\Theta_0 +
    2\Theta_2) - \frac{ck}{\mathcal{H}}\Theta_0^\prime}{(1+R)\tau^\prime + \frac{\mathcal{H}^\prime}{\mathcal{H}} -
    1}\\
    v_b^\prime &= \frac{1}{1+R} \left[-v_b - \frac{ck}{\mathcal{H}}\Psi + R(q +
    \frac{ck}{\mathcal{H}}(-\Theta_0 + 2\Theta_2) - \frac{ck}{\mathcal{H}}\Psi)\right]\\
    \Theta^\prime_1 &= \frac{1}{3} (q - v_b^\prime).
\end{align}
Note that $\Theta^\prime_0$ is still the same as given by equation \ref{eq:tp1}.  In the tight coupling regime the higher order multipoles are given by their initial as shown in equations \ref{eq:t2init} and \ref{eq:tellinit}.
\section{Results}
\label{sec:results}


\bibliography{ref.bib}
\bibliographystyle{aasjournal}
%\begin{thebibliography}{}
%\end{thebibliography}
\end{document}

% End of file `sample62.tex'.
